### 2022-01-28 chapter1/chapter.tex
same draft

%!TEX root = ../thesis.tex
\chapter{Introduction \& Preliminaries \label{ch:intro}} 
% THE ROUGH IDEA & CONTEXT
% 	THE CONTEXT: FLUID-FLUID QUEUES
%	PDEs 
%	NON-NEGATIVITY & CONSERVATION 
% MATHEMATICAL PRELIMINARIES
A fluid queue is a two-dimensional stochastic process \(\{\bs X(t)\} = \{(X(t),\varphi(t))\}_{t\geq0}\). The phase process, also known as the driving process, \(\{\varphi(t)\}_{t\geq0}\), is a continuous-time Markov chain (CTMC). The level process, \(\{X(t)\}_{t\geq0}\), is a real-valued, continuous, and piecewise linear.

@@

In the context of approximating fluid queues, one advantage of the QBD discretisation and, equivalently, DG schemes with constant basis functions, is that they guarantee probabilities computed from the approximation are positive \citep[Section 3.3]{koltai2011}, see also \citep{nodalDGBook} and references therein. One justification for the positivity preserving property is from the interpretation of the discretisation as a stochastic process thereby ensuring positivity. For higher order DG schemes there is no such interpretation. Moreover, higher-order DG approximation schemes may produce negative, or highly oscillatory solutions, particularly when discontinuities or steep gradients are present. Methods to navigate the problem of negative or highly oscillatory solutions have been developed, such as slope limiters, and filtering (see \citep[Section 6.5]{nodalDGBook} and references therein). Slope limiting effectively alters the discretised operator in regions where oscillations are detected and lowers the order of the approximation in theses regions. Filtering is a post-hoc method which looks to recover an accurate solution, given an oscillatory approximation.

\section{Fluid queues}
A fluid queue is a two-dimensional stochastic process \(\{\bs X(t)\} = \{(X(t),\varphi(t))\}_{t\geq0}\) where \(\{\varphi(t)\}_{t\geq0}\) is known as the phase or driving process, and \(\{X(t)\}_{t\geq0}\) is known as the level process or buffer. The phase process \(\{\varphi(t)\}_{t\geq0}\), is an irreducible continuous-time Markov chain (CTMC) with finite state space, which we we assume to be \(\mathcal S=\{1,2,\dots,N\}\) without loss of generality, and infinitesimal generator \(\bs T= [T_{ij}]_{i,j\in\mathcal S}\). We assume that \(\bs T\) is \emph{conservative}. Associated with states \(i\in\mathcal S\) are real-valued \emph{rates} \(c_i\in\mathbb R\).

@@

Let \(\bs f(x,t) = \vligne{f_i(x,t)}_{i\in\calS}\) be a row-vector function where \(f_i(x,t)\) is the density of \(\mathbb P(X(t)\leq x, \varphi(t) = i)\), assuming it exists. When a differentiable density exists, the system of partial differential equation which describes the evolution of the densities \(\bs f(x,t)\) is 
\begin{equation}
	\cfrac{\partial}{\partial t} \bs f(x,t) = \bs f(x,t)\bs T - \cfrac{\partial}{\partial x}\bs f(x,t)\bs C.\label{eqn: pde}
\end{equation}
The initial condition is the initial distribution of the fluid queue, \(f_i(x,0)\). Often a differentiable density function does not exist and therefore the partial differential equation (\ref{eqn: pde}) is not well-defined. For example, for a fluid queue with a regulated boundary, if the initial distribution of the fluid queue is a point mass at any point \(x_0\geq 0\) and in phase \(i\in\calS_+\cup\calS_0\), then a density function \(f_i(x,t)\) will not exist for any finite \(t\). Specifically, a point mass will persist along the ray \(x_0+c_it\), \(t\geq 0\). In such situations, it is the \emph{weak solution} to (\ref{eqn: pde}) that we seek. A weak solution satisfies
\begin{equation}
	-\int_x\int_t \bs f(x,t)\cfrac{\partial}{\partial t} \bs \psi(x,t)\wrt t \wrt x= \int_{x}\int_t\bs f(x,t)\bs T\bs \psi(x,t)\wrt t \wrt x + \int_x\int_t\bs f(x,t)\bs C \cfrac{\partial}{\partial x} \bs psi(x,t)\wrt t \wrt x,\label{eqn: weak pde}
\end{equation}
for every vector of test functions, \(\bs \psi(x,t)\), wich are smooth and have compact support.

@@

\section{Matrix-exponential distributions}
Here we recount some facts about matrix exponential distributions. See \citep{MEinAP} for a more detailed exposition. A random variable, \(Z\), is said to have a matrix-exponential distribution if it has a distribution function of the form \(1-\bs \alpha e^{\bs Sx}(-\bs S)^{-1}\bs s\), where \(\bs \alpha\) is a \(1\times p\) \emph{initial vector}, \(\bs S\) a \(p\times p\) matrix, and \(\bs s\) a \(p\times 1\) \emph{closing vector}, and \(\displaystyle e^{\bs S x} := \sum_{n=0}^\infty \cfrac{\left(\bs Sx\right)^n}{n!}\) is the matrix exponential. The density function of \(Z\) is given by \(f_Z(x) = \bs\alpha e^{\bs S x}\bs s\). The only restrictions on the parameters \((\bs \alpha, \bs S, \bs s)\) are that \(\bs \alpha e^{\bs S x} \bs s\) be a valid density function, i.e.~\(\bs \alpha e^{\bs S x} \bs s\geq 0,\) for all \(x\geq 0\) and \(\lim_{x\to\infty} 1-\bs \alpha e^{\bs Sx}(-\bs S)^{-1}\bs s = 1.\) There is the possibility of an \emph{atom} (a point mass) at 0, but here we do not consider this possibility. These condition that \(\bs \alpha e^{\bs Sx}\bs s\) be a valid density imposes some properties on representations \((\bs \alpha, \bs S, \bs s)\). However, in general there is no way to determine whether, given a triplet \((\bs \alpha, \bs S, \bs s)\) whether it is a representation of a matrix-exponential distribution, or not. Nonetheless, some properties of a triplet \((\bs \alpha, \bs S, \bs s)\) are known, such as the following, which is used in the characterisation of QBD-RAPs. 
\begin{thm}[Theorem 4.1.3, \cite{MEinAP}]
	The density function of a matrix-exponential distribution with representation \((\bs\alpha, \bs S, \bs s)\) can be expressed in terms of real-valued constants as 
	\begin{align}
		\psi(x)&=\sum_{j=1}^{m_1} \sum_{k=1}^{p_j} c_{jk} \cfrac{x^{k-1}}{(k-1)!} e^{\mu_jx} + \sum_{j=1}^{m+2}\sum_{k=1}^{q_j} d_{jk}\cfrac{x^{k-1}}{(k-1)!}e^{\eta_jx}\cos(\sigma_jx) \nonumber 
		\\&\quad{}+ \sum_{j=1}^{m_2}\sum_{k=1}^{q_j}e_{jk}\cfrac{x^{k-1}}{(k-1)!}e^{\eta_jx}\sin(\sigma_jx), \label{eqn: spec}
	\end{align}
	for integers \(m_1,\,m_2,\,p_j,\) and \(q_j\) and some real constants \(c_{jk},\,d_{jk},\,e_{jk},\,\mu_j,\,\eta_j,\) and \(\sigma_j\). Here \(\mu_j,\, j=1,\dots,m_1\) are the real eigenvalues of \(\bs S\), while \(\eta_j+i\sigma_j, \, \eta_j-i\sigma_j,\, j=1,\dots,m_2\) denote its complex eigenvalues, which come in conjugate pairs. Thus \(m_1+2m_2\) is the total number of eigenvalues, while the dimension of the representation is given by \(\displaystyle p=\sum_{j=1}^{m+1}p_j + 2\sum_{j=1}^{m_2}q_j\). 
\end{thm}
\begin{thm}[Theorem 4.1.4, \cite{MEinAP}]\label{thm: 4.1.4}
	Consider the nonvanishing terms of the matrix exponential density (\ref{eqn: spec}), \emph{i.e.}, the terms for which \(c_{jk}=0,\,d_{jk} =0,\) or \(e_{jk}=0\). Among the corresponding eigenvalues \(\lambda_j\), there is a real dominating eigenvalue \(\kappa\), say. That is, \(\kappa\) is real, \(\kappa \geq Re(\lambda_j)\) for all \(j\), and the multiplicity of \(\kappa\) is at least the multiplicity of every other eigenvalue with real part \(\kappa\).
\end{thm}
\begin{cor}[Corollary 4.1.5, \cite{MEinAP}]
If \((\bs \alpha, \bs S, \bs s)\) is a representation for a matrix-exponential distribution, then \(\bs S\) has a real dominating eigenvalues.
\end{cor}
\begin{thm}[Theorem 4.1.6, \cite{MEinAP}]
	Let \(Z\) be a matrix-exponentially distributed random variable with density (\ref{eqn: spec}). Then the dominant real eigenvalue \(\kappa\) of Theorem~\ref{thm: 4.1.4} is strictly negative. 
\end{thm}
We define \(dev(\bs S)\) to be the real dominating eigenvalue of \(\bs S\), that is \(dev(\bs S)=\kappa\) in Theorem~\ref{thm: 4.1.4}.

>>>>

%!TEX root = ../thesis.tex
\chapter{Introduction \& Preliminaries \label{ch:intro}} 
% intro to intro
% contextual background
	% Fluid queues 
		% definition
		% applications 
		% existing results 
	% Fluid-fluid queues
		% definition 
		% basic analysis in terms of generators
		% the need for approximation and how the old work doesn't fit here
	% The DG method
		% pertition into cells
		% project on to a basis
		% problems: oscillations
		% soutions and why they dont fit
		% observation about constant-basis methods
		% this thesis asks the question: motivated by the constant basis being a probability model, can we dream up a more accurate probability model to approximate a fluid queue?
	% On the structure of the constant basis/uniformisation method (QBD)
	% least variable is PH, so what about MEs?
% structure of the thesis
	% rest of chapter 1 gives mathematical preliminaries
	% chapter 2 explains the DG method, problems/oscillations, slope limiting, and appllication to SFFMs
	% inspired by the structure of the order-1 scheme, and to solve the negativity problem, chapter 3 introduces a new approximation method; The QBD-RAP. In this chapter we derive the intuitively of the model and explain it's dynamics.
	% we then move to convergence, with chapter 4 proving a convergence of the QBD-RAP up to the first hitting time of the fluid queue on the edges of an interval. technical results and extensions are left to the appendix: certain properties of closing vectors, convergence without ephemeral phases, and certain matrix algebraic manipulations. 
	% chapter 5 stitches together the results of chapter 4 to prove a global result about convergence
	% chapter 6 investigate, numerically, the performance of the DG, DG with limiter, uniformisation and QBD-RAP schemes. 
	% chapter 7  makes concluding remarks.
% mathematical preliminaries 
	% CTMCs
	% fluid queues
	% fluid-fluid queues and operator-analytic expressions (all of them)
		% the equations of bo2014
		% meets the partition of the DG/QBD-RAP, and reconstructing the Fil sets 
    % why we cant solve the equations, what we need to approximate to solve it, i.e. B, R, then D, Psi
	% projections
	% phase-type distributions 
		% definition
		% least variable property
	% matrix exponentials
		% definition 
		% properties 
		% verification that parameters give an ME
		% CMEs
	% QBD-RAPs, then QBDs as a special case
		% orbit processes, their interpretation for PH and how they differ for MEs
	% convergence theorems 
		% DCT 
		% Continuity of Laplace transforms

% THE ROUGH IDEA & CONTEXT
% 	THE CONTEXT: FLUID-FLUID QUEUES
%	PDEs 
%	NON-NEGATIVITY & CONSERVATION 
% MATHEMATICAL PRELIMINARIES
A fluid queue is a two-dimensional stochastic process \(\{\bs X(t)\} = \{(X(t),\varphi(t))\}_{t\geq0}\). The phase process, also known as the driving process, \(\{\varphi(t)\}_{t\geq0}\), is a continuous-time Markov chain (CTMC). The level process, \(\{X(t)\}_{t\geq0}\), is a real-valued, continuous, and piecewise linear.

@@

In the context of approximating fluid queues, one advantage of the QBD discretisation and, equivalently, DG schemes with constant basis functions, is that they guarantee probabilities computed from the approximation are positive \citep[Section 3.3]{koltai2011}, see also \citep{nodalDGBook} and references therein. One justification for the positivity preserving property is from the interpretation of the discretisation as a stochastic process thereby ensuring positivity. For higher order DG schemes there is no such interpretation. Moreover, higher-order DG approximation schemes may produce negative, or highly oscillatory solutions, particularly when discontinuities or steep gradients are present. Methods to navigate the problem of negative or highly oscillatory solutions have been developed, such as slope limiters, and filtering (see \citep[Section 6.5]{nodalDGBook} and references therein). Slope limiting effectively alters the discretised operator in regions where oscillations are detected and lowers the order of the approximation in these regions. Filtering is a post-hoc method which looks to recover an accurate solution, given an oscillatory approximation.

\section{Mathematical Preliminaries}
\subsection{Fluid queues}
A fluid queue is a two-dimensional stochastic process \(\{\bs X(t)\} = \{(X(t),\varphi(t))\}_{t\geq0}\) where \(\{\varphi(t)\}_{t\geq0}\) is known as the phase or driving process, and \(\{X(t)\}_{t\geq0}\) is known as the level process or buffer. The phase process \(\{\varphi(t)\}_{t\geq0}\), is an irreducible continuous-time Markov chain (CTMC) with finite state space, which we we assume to be \(\mathcal S=\{1,2,\dots,N\}\) without loss of generality, and infinitesimal generator \(\bs T= [T_{ij}]_{i,j\in\mathcal S}\). We assume that \(\bs T\) is \emph{conservative}. Associated with states \(i\in\mathcal S\) are real-valued \emph{rates} \(c_i\in\mathbb R\).

@@

Let \(\bs f(x,t) = \vligne{f_i(x,t)}_{i\in\calS}\) be a row-vector function where \(f_i(x,t)\) is the density of \(\mathbb P(X(t)\leq x, \varphi(t) = i)\), assuming it exists. When a differentiable density exists, the system of partial differential equation which describes the evolution of the densities \(\bs f(x,t)\) is 
\begin{equation}
	\cfrac{\partial}{\partial t} \bs f(x,t) = \bs f(x,t)\bs T - \cfrac{\partial}{\partial x}\bs f(x,t)\bs C.\label{eqn: pde}
\end{equation}
The initial condition is the initial distribution of the fluid queue, \(f_i(x,0)\). Often a differentiable density function does not exist and therefore the partial differential equation (\ref{eqn: pde}) is not well-defined. For example, for a fluid queue with a regulated boundary, if the initial distribution of the fluid queue is a point mass at any point \(x_0\geq 0\) and in phase \(i\in\calS_+\cup\calS_0\), then a density function \(f_i(x,t)\) will not exist for any finite \(t\). Specifically, a point mass will persist along the ray \(x_0+c_it\), \(t\geq 0\). In such situations, it is the \emph{weak solution} to (\ref{eqn: pde}) that we seek. A weak solution satisfies
\begin{equation}
	-\int_x\int_t \bs f(x,t)\cfrac{\partial}{\partial t} \bs \psi(x,t)\wrt t \wrt x= \int_{x}\int_t\bs f(x,t)\bs T\bs \psi(x,t)\wrt t \wrt x + \int_x\int_t\bs f(x,t)\bs C \cfrac{\partial}{\partial x} \bs \psi(x,t)\wrt t \wrt x,\label{eqn: weak pde}
\end{equation}
for every vector of test functions, \(\bs \psi(x,t)\), wich are smooth and have compact support.

@@

\subsection{Fluid-fluid Queues}
	\label{sec:prelim}
	\begin{center}
		\begin{minipage}{0.8\textwidth}
			\textit{This subsection has been taken from Sections~2 and 3 of \cite{blnos2022} with only minor changes, such as notations, so that this chapter is consistent with the rest of the thesis. I am a co-author of the paper \cite{blnos2022}. The conceptualisation of \cite{blnos2022} was originally by Vikram Sunkara, Nigel Bean and Giang Nguyen, and the original coding was done by Vikram Sunkara. I made significant contributions to Section~3 of the paper, expressing the operator-theoretic expressions to use the same partition as the approximation scheme. I contributed Sections~4.4 and 5.1. I extended the numerical experiments in Section~6 to higher orders and made all the plots in Section~6. Appendix~A is also my original work. I did a significant proportion of the writing of the manuscript and addressed the reviewers comments and also developed code for the numerical experiments.
			}
		\end{minipage}
		\end{center}

An unbounded stochastic fluid-fluid model (SFFM) is a Markov process $\{(\widehat{X}_t, \widehat{Y}_t, \varphi(t)),t \geq 0\},$ where the phase $\{\varphi(t)\}$ is a continuous-time Markov chain on a finite state space $\mathcal{S}$;  $\{\widehat{X}_t\}$ is the first fluid, which varies linearly at rate~$c_{\varphi(t)}$
% 
\begin{align*} 
	\widehat X(t) := \widehat X(0) + \int_0^t c_{\varphi(s)} \wrt s;
\end{align*} 
% 
and $\widehat{Y}_t$ is the second fluid, which varies at rate $r_{\varphi(t)}(\widehat{X}_t)$:
% 
\begin{align*} 
	\widehat{Y}_t := \widehat Y(0)) + \int_0^t r_{\varphi(s)}(\widehat{X}_s) \wrt s.
\end{align*} 
% 
Regulated boundaries may also be included for both fluids. To distinguish between unbounded and bounded processes, we use the notations \(\widehat X(t)\) and \(\widehat Y(t)\) to denote unbounded processes, and \(X(t)\) and \(Y(t)\) to denote fluid levels with a regulated lower boundary at 0.

As classic fluid processes, $\{(\widehat X(t), \varphi(t)),t \geq 0\}$, or unbounded analogues, are used extensively in many areas, such as insurance and environmental modelling, it is clear that stochastic fluid-fluid models have an even wider range of applicability.

We assume that $X(t), Y(t) \in [0,\infty)$ and that there is a regulated boundary at level $0$ for both buffers: 
% 
	\begin{align*} 
		\frac{\wrt}{\wrt t} X(t) & := \max\{0, c_i\} \quad \mbox{if } X(t) = 0 \mbox{ and } \varphi(t) = i, \\
          	\frac{\wrt}{\wrt t} Y(t) & := \max\{0, r_i(x)\} \quad \mbox{if } Y(t) = 0,\,X(t) = x \mbox{ and } \varphi(t) = i, 	
	\end{align*} 
	% 
for $i \in \mathcal{S} = \{1,...,N_\calS\}$. Let $T$ be the irreducible generator for the finite-state Markov chain $\{\varphi(t)\}$. 
%  on $\mathcal{S} = \{1, \ldots, L\}$. 
We denote by $\bs R(x) := \diag(r_i(x))_{i \in \mathcal{S}}$ the diagonal fluid-rate matrix of functions for $\{Y(t)\}$.

\begin{rem}[Notation]\label{remark: notation 1}For future reference we require some notation regarding the elements of the model introduced above. We use the notation \(\bs u = (u_h)_{h\in \mathcal H}\) to denote a row-vector, \(\bs u\), defined by its elements, \(u_h\), indexed by \(h\in\mathcal H\), where \(\mathcal H\) is some countable index set. Similarly, \(\bs u = (\bs u_h)_{h\in\mathcal H}\), is a row-vector defined by a collection of row-vectors \(\bs u_h\). The notation \(\bs u_m=(u_h)_{h\in\mathcal H_m}\) refers to the vector containing the subset of elements corresponding to \(\mathcal H_m\subseteq \mathcal H\). When the index set is empty, the resulting vector \(\bs u_m\) is a vector of dimension 0. In cases when there are two indices, we order the elements of the vector according to the first index, then the second; i.e.~\(\bs u = (u_{g}^h)_{g\in\mathcal G,h\in\mathcal H} = ((u_g^h)_{g\in\mathcal G})_{h\in\mathcal H}\). Here we use the convention that for a vector \(\bs u=(u)_{h\in \mathcal H}\) where the elements \(u\) do not depend on the index \(h\) and \(H\) is some index set, then we repeat \(u\) \(h\)-times; i.e.~\(\bs u = (u)_{h\in \mathcal H} =\underbrace{(u,\dots,u)}_{h-\mbox{times}}\). The notation \(\bs U = [u_{gh}]_{g\in\mathcal G, h\in\mathcal H}\) is used to denote a matrix defined by its elements, or sub-blocks, \(u_{gh}\). 
\end{rem}

Let \(\mathcal S_-=\{i\in\calS\mid c_i<0\},\, \mathcal S_+=\{i\in\calS\mid c_i>0\},\, \mathcal S_0=\{i\in\calS\mid c_i=0\},\, \mathcal S_\nabla=\{i\in\calS\mid c_i\leq0\},\, \mathcal S_\Delta=\{i\in\calS\mid c_i\geq0\}\).

For the remainder of this section, we summarise the findings of~\citep{bo2014} on the joint stationary distribution of $\{(X(t), Y(t), \varphi(t)),t \geq 0\}$. For each Markovian state $i \in \mathcal{S}$, we partition the state space of \(X(t)\), \([0,\infty)\), according to the rates of change $r_i(\cdot)$ for the second fluid $\{Y(t)\}$: $[0,\infty) := \mathcal{F}^{+}_i \cup \mathcal{F}^{-}_i \cup \mathcal{F}^{0}_i,$  
where 
% 
		\begin{align} 
%				\label{eqn:Fi1}
			\mathcal{F}^{+}_i & := \{u \in \mathcal{F} : r_i(u) > 0\},  \;
%				\label{eqn:Fi2}
			\mathcal{F}^{-}_i :=  \{u \in \mathcal{F}:  r_i(u) < 0\}, \;
%				\label{eqn:Fi3}
			\mathcal{F}^{0}_i := \{u \in \mathcal{F}: r_i(u) = 0\}.\label{eqn:fil}
		\end{align} 
% 
For all $i \in \mathcal{S}$, the functions $r_i(\cdot)$ are assumed to be sufficiently well-behaved that $\mathcal{F}^{m}_i$, $m \in \{+, -, 0\}$, is a finite union of intervals and isolated points.

We assume that the process $\{(X(t), Y(t), \varphi(t)),t\geq 0\}$ is positive recurrent, in order to guarantee the existence of the joint stationary density. Define stationary operators 
\begin{align} 
		\label{eqn:jointpi} 
	\overline \bbpi_i(y)(\mathcal{A}) & := \lim_{t \rightarrow \infty} \frac{\partial}{\partial y} \mathbb{P}\left(X(t) \in \mathcal{A}, Y(t) \leq y, \varphi(t) = i\right), \quad y>0,\\
% 
		\label{eqn:jointmass}
		\overline{\mathbb p}_i(\mathcal{A}) & := \lim_{t \rightarrow \infty}  \mathbb{P}(X(t) \in \mathcal{A}, Y(t) = 0, \varphi(t) = i]),
	\end{align} 
	where $\mathcal{A} \subset [0,\infty)$.
Then let $\overline{{\bbpi}}(y) = (\overline \bbpi_i(y))_{i \in \mathcal{S}}$ be a vector containing the joint stationary density operators and $\overline{\mathbb{p}} = (\overline{\mathbb p}_i)_{i \in \mathcal{S}}$ a vector containing the joint stationary mass operators.

The determination of $\overline{{\bbpi}}(y)$ involves two important matrices of operators, $\overline{\mathbb B}$ and $\overline{\mathbb \Psi}$. The operator \(\overline{\mathbb B}\) is the \textit{infinitesimal generator} of the process \(\{(X(t),\varphi(t))\}\). %Intuitively, for a set $\mathcal{A} \subset \mathcal{F}$ and a measure vector $\boldsymbol{\mu} = (\mu_i)_{i \in \mathcal{S}}$, $\bs{\mu}V(t)(\mathcal{A})$, where \(V(t)\) is a semigroup with infinitesimal generator \(B\), gives the conditional probability of $X(t) \in \mathcal{A}$, 
The operator \(\overline{\mathbb \Psi}\) is such that $\boldsymbol{\mu} \overline{\mathbb \Psi} (\mathcal{A})$ is the conditional probability of $\{Y(t)\}$ returning to level zero and doing so when $X(t) \in \mathcal{A}$, given that the initial distribution is $\boldsymbol{\mu}$.

\subsubsection{Matrix $\overline{\mathbb B}$ of Operators}
\label{subsec:B_operators}
Since \(\{(X(t),\varphi(t)),t\geq 0\}\) is a Markov process, the evolution of probability can be described by a semigroup. Let $\mathcal{M}(\mathcal{S} \times \mathbb{R}_{+})$ be the set of integrable complex-valued Borel measures on the Borel $\sigma$-algebra $\mathcal{B}_{\mathcal{S} \times \mathbb{R}_{+}}$. For $\bs{\mu} \in \mathcal{M}(\mathcal{S} \times \mathbb{R}_{+})$, we can write \(\bs{\mu} = (\mu_i)_{i \in \mathcal{S}}\). The measures \(\mu_i(\cdot)\) represent an initial distribution, \(\mu_i(\cdot) = \mathbb P(X(0))\in\cdot, \varphi(0) = i)\). 
Let \(\{\overline{\mathbb V}(t)\}_{t\geq 0},\, \overline{\mathbb V}(t):\mathcal{M}(\mathcal{S} \times \mathbb{R}_{+})\mapsto \mathcal{M}(\mathcal{S} \times \mathbb{R}_{+})\) be the semigroup describing the evolution of probability for \(\{(X(t),\varphi(t)),t\geq 0\}\) structured as a matrix of operators, \(\left[ \overline{\mathbb V}(t)\right]_{ij}= \overline{\mathbb V}_{ij}(t)\) where, 
\[\mu_i \overline{ \mathbb V}_{ij}(t)(\mathcal A) = \int_{x\in[0,\infty)} \wrt \mu_i(x)\mathbb P (X(t) \in\mathcal A,\varphi(t) = j \mid X(0)) = x, \varphi(0) = i).\]
Intuitively, the operator \( \overline{\mathbb V}(t)\) maps an initial measure \(\boldsymbol \mu\) on \((X(0),\varphi(0))\) to the measure \(\mathbb P(X(t)\in \mathcal A, \varphi(t)=j)=:\mu_j(t)(\mathcal A)\). 
The matrix of operators \( \overline{\mathbb B}:=[\overline{\mathbb B}_{ij}]_{i,j\in\mathcal S}\) is the \textit{infinitesimal generator} of the semigroup \(\{ \overline{\mathbb V}(t)\}\) defined by 
\[ \overline{\mathbb B} =\left.\cfrac{\wrt}{\wrt t} \overline{\mathbb V}(t)\right|_{t=0},\]
with domain the set of measures for which this limit exists. Specifically, the domain of \(\overline{\mathbb B}\) is the set of measures, \(\bs \mu = (\mu_i)_{i\in\calS}\), for which each \(\mu_i\) admits an absolutely continuous density on \((0,\infty)\), and can have a point mass at \(0\) if \(i\in\mathcal S_\nabla\); call this set of measures \(\mathcal M_0\). That is, the measures \(\mu_i,\, i\in\calS,\) such that \((\mu_i)_{i\in\calS}\in \mathcal M_0\) are absolutely continuous on \((0,\infty)\) and may have point masses at \(0\) if \(i\in\calS_\nabla\). The measure \(\mu_i\) cannot have a point mass at 0 if \(i\notin \mathcal S_\nabla\). In the sequel we write \(v_i(x),\, x>0\), as the density of \(\mu_i\) and \(q_i\) as the point mass of \(\mu_i\) at \(x=0\) (if such a point mass exists).

To use the operators \(\{ \overline{\mathbb  V}(t)\}\) and \( \overline{\mathbb  B}\) to analyse the fluid-fluid model, \cite{bo2014} explicitly track when \((X(t),\varphi(t))\in(\mathcal F_i^m,i)\) for \(i\in\mathcal S,\, m \in \{+,-,0\}\) by partitioning the operators \( \overline{\mathbb  V}(t)\) and \( \overline{\mathbb  B}\) into \( \overline{\mathbb  V}_{ij}^{m n}\) and \( \overline{\mathbb  B}_{ij}^{m n}\), for \(i,j\in\mathcal S,\, m,n\in \{+,-,0\}\), where
\[\left.\mu_i\right|_{\calF_i^m}  \overline{\mathbb  V}_{ij}^{m n}(t)(\mathcal A) := \int_{x\in[0,\infty)} \left. \wrt \mu_i\right|_{\calF_i^m}(x)\mathbb P (X(t) \in\mathcal A\cap \calF_j^n,\varphi(t) = j \mid X(0) = x, \varphi(0) = i),\]
and \(\left.\mu_i\right|_{E}\) is the restriction of \(\mu_i\) to \(E\). Similarly, for \( \overline{\mathbb  B}_{ij}^{m n},\) \(i,j\in\mathcal S,\, m,n\in \{+,-,0\}\).

We claim that numerical schemes are needed to approximate the analytic operator equations introduced in \cite{bo2014}. The DG scheme we choose to use here works by first partitioning the state space of the fluid level, \(\{X(t)\}\), into a collection of intervals, \(\calD_k=[x_k,x_{k+1}]\) then, on each interval, the operator \(\overline{\mathbb B}\) is projected onto a basis of polynomials. So, to help elucidate the connection between the operators \(\{ \overline{\mathbb  V}(t)\}\), \( \overline{\mathbb  B}\) and their DG approximation counterparts, we take a slightly different approach to partitioning these operators than that taken in \cite{bo2014}. Rather than partition according to the sets \(\calF_i^m,\,i\in\calS,\,m\in\{+,-,0\}\), we use the same partition as that in the construction of the DG scheme. By doing so, we can directly correspond elements of the partitioned operators to their approximation counterparts. Since the partition used to construct the DG scheme is finer, then we can always reconstruct the partition in terms of the sets \(\calF_i^m,\,i\in\calS,\,m\in\{+,-,0\}\).

Let us first partition the space \([0,\infty)\) into \(\calD_\nabla = \{0\}\), and non-trivial intervals \(\calD_k=[x_k,x_{k+1}]\setminus \{0\},\) with \(x_1=0,\, x_k<x_{k+1},\,k=1,2,...\). The symbol \(\nabla\) is used to refer to sets and quantities which are relevant to boundary at \(x=0\). For $\bs{\mu} \in \mathcal{M}_0(\mathcal{S} \times \mathbb{R}_{+})$ we write \(\bs{\mu} = (\mu^{k}_i)_{i \in \mathcal{S},k\in\{\nabla,1,2,...\}},\) where $\mu^{k}_i(\cdot) = \mu_i(\cdot \cap \calD_k),\,k=\nabla,1,2,\dots.$ We also have densities, \(v_i^k(x),\,x>0\), associated with each measure, \(\mu_i^k\). For \( i,j\in\calS,\,k,\ell\in\{\nabla,1,2,\dots\}\) define the operators 
\[\mu_i^k \mathbb V_{ij}^{k \ell}(t)(\mathcal A) := \int_{x\in\calD_k} \wrt \mu_i^k(x)\mathbb P (X(t) \in\mathcal A\cap \calD_\ell,\varphi(t) = j \mid X(0) = x, \varphi(0) = i),\]
and the matrices of operators \(\mathbb V^{k\ell}(t) := \vligne{\mathbb V_{ij}^{k\ell}(t)}_{i,j\in\mathcal S},\,k,\ell\in\{\nabla,1,2,\dots\}\) and write 
\[\mathbb V(t) = \left[
	\begin{array}{llll}
		\mathbb V^{\nabla,\nabla}(t)&\mathbb V^{\nabla,1}(t)& \mathbb V^{\nabla,2}(t) &\hdots \\
		\mathbb V^{1,\nabla}(t)&\mathbb V^{1,1}(t)&\mathbb V^{1,2}(t)&\hdots\\
		\mathbb V^{2,\nabla}(t)&\mathbb V^{2,1}(t)&\mathbb V^{2,2}(t)&\hdots\\
		\vdots & \vdots & \vdots & \ddots 
%		&\ddots &&&&\reflectbox{\(\ddots\)}\\
%		&&\mathbb V^{k-1,k-1}(t) & \mathbb V^{k-1,k}(t) & \mathbb V^{k-1,k+1}(t)& \\
%		&&\mathbb V^{k,k-1}(t) & \mathbb V^{k,k}(t) & \mathbb V^{k,k+1}(t)& \\
%		&&\mathbb V^{k+1,k-1}(t) & \mathbb V^{k+1,k}(t) & \mathbb V^{k+1,k+1}(t)& \\
%		&\reflectbox{\(\ddots\)}&&&&\ddots
	\end{array}\right].\]
Now define \(\mathbb B=\left.\cfrac{\wrt}{\wrt t}\mathbb V(t)\right|_{t=0}\) as the infinitesimal generator of \(\{\mathbb V(t)\}\), resulting in the tridiagonal matrix of operators 
\[\mathbb B(t) = \left[
	\begin{array}{llll}
		\mathbb B^{\nabla,\nabla}(t)&\mathbb B^{\nabla,1}(t)& & \\
		\mathbb B^{1,\nabla}(t)&\mathbb B^{1,1}(t)&\mathbb B^{1,2}(t)&\\
		&\mathbb B^{2,1}(t)&\mathbb B^{2,2}(t)&\ddots\\
		& & \ddots & \ddots 
	\end{array}\right],\]
%\begin{align*}
%\mathbb B &= \left[\begin{array}{lllll}
%		\ddots &&&&\\
%		\mathbb B^{k-2,k-1} &\mathbb B^{k-1,k-1} & \mathbb B^{k-1,k} & & \\
%		&\mathbb B^{k,k-1} & \mathbb B^{k,k} & \mathbb B^{k,k+1}& \\
%		& & \mathbb B^{k+1,k} & \mathbb B^{k+1,k+1}&\mathbb B^{k+1,k+2} \\
%		&&&&\ddots
%	\end{array}\right],
%\end{align*}
where the blocks \(\mathbb B^{k\ell}:= \vligne{\mathbb B_{ij}^{k\ell}(t)}_{i,j\in\mathcal S},\,k,\ell\in\{\nabla,1,2,\dots\}\). 
The tridiagonal structure arises since, for \(|k-\ell|\geq2\) (where we take \(\nabla = 0\) if it appears in the differences) it is impossible for \(\{X(t)\}\) to move from \(\mathcal D_k\) to \(\mathcal D_\ell\) in an infinitesimal amount of time.

\begin{rem}[More notation]\label{remark: notation 2} We use a blackboard bold font with an overline above the character (e.g.~\(\overline{\mathbb B}\) and \(\overline{\mathbb V}(t)\)) to represent theoretical operators derived in \citep{bo2014} which are constructed using the partition in (\ref{eqn:fil}). The operators denoted with an overline play a minor role in the introductory sections of this paper, but do not appear again. We use a blackboard font sans overline (e.g.~\(\mathbb V(t)\) and \(\mathbb B\)) to represent the same operators but which are constructed with the finer partition defined by \(\mathcal D_k,\,k=\nabla,1,2,\dots\). We use the letters \(i,j\in\mathcal S\) to represent states of the phase process, letters \(m,n,\in\mathcal \{+,-,0\}\) to refer to the partition in terms of the sets in Equations (\ref{eqn:fil}), and the letters \(k,\ell\in\{\nabla,1,2,...\}\) to refer to the finer partition into sets \(\{\calD_k\}_k\). With a slight abuse of notation, whenever we use the dummy variables \(k,\ell\) we imply \(k,\ell\in\{\nabla,1,2,\dots\}\), the dummy variables \(m,n\) imply \(m,n\in\{+,-,0\}\) and the dummy variables \(i,j\) imply \(i,j\in\calS\). E.g.~\(\mathbb B_{ij}^{k\ell}\) means \(\mathbb B_{ij}^{k\ell},\, i,j\in\calS, k,\ell\in\{\nabla,1,2,\dots\}\) and \(\mathbb B_{ij}^{mn}\) means \(\mathbb B_{ij}^{mn},\, i,j\in\calS, m,n\in\{+,-,0\}\).
\end{rem}
